Multitask matrix completion for learning protein interactions across diseases

### 2016-11-22 paper.tex
adding eps figures
new stuff

\documentclass{bioinfo}
\copyrightyear{2015} \pubyear{2015}

%\usepackage[top=1.3in, bottom=1.3in, left=1in, right=1in]{geometry}
\usepackage{floatrow}
\usepackage{graphicx} % more modern
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{subfigure} 
\usepackage{natbib}
%\usepackage[footnotesize]{caption}
\usepackage{multirow}
\usepackage{url}

@@

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{Subject Section}

\title[Multitask matrix completion]{Multitask matrix completion for learning protein interactions across diseases}

\author[Kshirsagar \textit{et~al}.]{Meghana Kshirsagar$^{1}$, Keerthiram Murugesan$^{2}$, Jaime G. Carbonell$^{2}$ and Judith Klein-Seetharaman$^{3}$}
\address{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021\\
$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK
}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{\textbf{Motivation:}
Disease causing pathogens such as viruses, introduce their proteins into the host cells 
where they interact with the host's proteins enabling the virus to replicate inside the host.
These interactions between pathogen and host proteins are key to understanding infectious diseases.
Often multiple diseases involve phylogenetically related or biologically similar pathogens. Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}. \\
\textbf{Results:}
Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
We show how our model's parameters can be interpreted to reveal both general and specific
interaction-relevant characteristics of the viruses. \\
\textbf{Availability:}
Code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
\textbf{Contact:}
\href{mkshirsa@cs.cmu.edu}{mkshirsa@cs.cmu.edu}
}

@@

\begin{figure*}
\begin{floatrow}
\capbtabbox[0.68\textwidth]{%
\begin{footnotesize}
\begin{tabular}{c|r|r|r} 
\hline
Task $\rightarrow$ &{\textit{Influenza A} } &\textit{Hepatitis C} & {\textit{Ebola}}\\
& (28 strains) & (5 strains, many isolates)  & (3 strains) \\
number of HP PPIs (positives) & 848 & 981 & 90 \\\hline
number of non-homologous PPIs & 274 & 270 & 75 \\ \hline
\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
total \# of virus proteins in & 542 & 163 & 38 \\
UniProtKB across strains $^\dagger$ & & & \\ \hline
number of negatives & 84800 & 98100 & 9000 \\\hline
density of observed graph$^\ddagger$ (as \%) & .15 & .60 & .20 \\\hline
\multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
\multicolumn{4}{l}{\scriptsize{$\ddagger$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The denominator was computed }} \\
\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} considering all proteins from all strains of the virus and $\approx$ 100,000 human proteins.}}\\
\multicolumn{4}{l}{\scriptsize{$\dagger$ this is the number of `reviewed' proteins on UniProtKB from all strains in our data}} \\ \hline
\end{tabular}
\end{footnotesize}
}
{
\caption{Tasks and their sizes (a column corresponds to one bipartite graph between human proteins and the virus indicated in the column header). More importantly, each task represents several strains of one virus species. We pool together PPI from all strains as the number of interactions per strain is very small and not representative. PPI data from different strains often involves homologous proteins (virus proteins or/and human proteins) as these were curated from different publications/studies involving the virus. The second row gives a conservative estimate of the number of non-homologous PPI. These were obtained by computing BLAST sequence similarity (see Section \ref{sec:phisto} for details). All pathogens are single stranded RNA viruses. The last row shows that each of our graphs is extremely sparse.} 
\label{datasets}
}
\ffigbox[0.28\textwidth]{%
%trim option's parameter order: left bottom right top
%\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
\includegraphics[scale=0.34, trim = 0 0 0 0]{pca_original_feats.eps}
}
{
\caption{Principal component analysis (PCA) of virus proteins in the original feature space. The first two principal components are shown. Shape and color of the points indicates which virus that protein comes from.}%
\label{fig:pca1}
%\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
%\label{fig:hpppigraph}
}
\end{floatrow}
\end{figure*}

@@

\begin{figure*}
\begin{floatrow}
\capbtabbox[0.68\textwidth]{%
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|ccccccc}
\toprule
& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
MMTL & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
Sparse+low-rank & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
MTPL & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
IMC & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
BSL-MTL & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
}
{
\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset. The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data. We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100). The standard deviation is also shown. The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold. The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task. This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
\label{resultsTable}
}

\ffigbox[0.28\textwidth]{%
%%trim option's parameter order: left bottom right top
\includegraphics[scale=0.28, trim = 1.1cm 0.9cm 0 0]{pca_bslmtl_with_clusters.eps}
}
{
\caption{Principal component analysis (PCA) of virus proteins in the projected subspace. The first two principal components are shown. Shape of the points indicates which virus that protein comes from. We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
\label{fig:pca2}
}
\end{floatrow}
\end{figure*}

@@

\section{Results}
The PPI data we obtain from PHISTO has homologous PPI within a task. Note that the presence of sequence similarity across virus proteins is a strong criterion for computational methods to work, however orthologs or paralogs of proteins within a task lead to redundancy in the data and some readers may find it harder to judge the contribution of the methods. We therefore present results on three sets: 
\begin{enumerate}[a]
\item the original PPIs from PHISTO. AUC-PR is shown in Table \ref{resultsTable}
\item \textbf{Non-homologous partition}: after removal of all homologous PPI from the test data. Here we ensure that the test data does not contain homologs of PPI observed in the training data (this ensures there are no redundant examples). Further, every set of homologous PPIs in the test data is replaced with a single PPI. A BLAST sequence similarity cut-off of $1e^{-5}$ was used to find homologs. The results are in Table \ref{nohomresultsTable}.
\item \textbf{Homologous PPI partition} : only homologous PPI part of the test data. This is essentially the set of PPI removed in the set-(b) above. Table \ref{homresultsTable} shows these results. 
\end{enumerate}
Note that the partitions in (b) and (c) together comprise our test data. Also, the negatives in the test data are randomly split into the `no-homolog' and 'homolog' partitions while maintaining the number of negatives in each partition to be 100 times the number of positives.

@@

\begin{table}[!h]
\caption{AUC-PR on the \textbf{non-homologous PPI within each task}, in the 10\% setting. To get an estimate of the relative size of the non-homologous partition for each task, see Table \ref{datasets}. Only the most representative baselines are retained for this experiment. Our model performs better than the others on the non-homologous part of the test data, as it captures non-linear similarities among the interactions within a task and across tasks very well.}
\label{nohomresultsTable}
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|ccc}
\toprule
\multicolumn{4}{c}{\textbf{10 \% training, 90\% test on Non-homologous PPI}} \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & 0.207 $\pm$.07 & 0.113 $\pm$.01 & 0.109 $\pm$ .01 \\
STL  & 0.210 $\pm$.14 & 0.474 $\pm$ .06 & 0.302 $\pm$.03 \\
Sparse+LR & 0.216 $\pm$.13 & 0.482 $\pm$.05 & 0.312 $\pm$.04 \\ 
BSL-MTL & \textbf{0.295} $\pm$.10 & \textbf{0.544} $\pm$.06 & \textbf{0.505} $\pm$.05 \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

\begin{table}[!h]
\caption{AUC-PR on the \textbf{homologous PPI} within each task, in the 10\% setting. The Homolog baseline outperforms all on the \textit{Ebola} task and BSL-MTL beats all other methods on the other tasks. The Homolog method has access to a wider virus-human PPI graph involving several other viruses and is able to infer homologous PPI for \textit{Ebola} better (note that some of the homologous PPI are not redundant examples but involve proteins with very small regions of sequence similarity.}
\label{homresultsTable}
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|ccc}
\toprule
\multicolumn{4}{c}{\textbf{10 \% training, test only Homologous PPI }} \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & \textbf{0.399} $\pm$.07 & 0.161 $\pm$.01 & 0.180 $\pm$ .01 \\
STL   & 0.192 $\pm$.15 & 0.504 $\pm$.10 & 0.305 $\pm$.03 \\
Sparse+LR & 0.184 $\pm$.13 & 0.512 $\pm$.10 & 0.315 $\pm$.04 \\ 
BSL-MTL & 0.130 $\pm$.10 & \textbf{0.815} $\pm$.02 & \textbf{0.475} $\pm$.02 \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

\begin{figure}
\begin{tabular}{c}
%\begin{minipage}{0.3\textwidth}
%\begin{center}
%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
%\end{center}
%\end{minipage}
%&
\begin{minipage}{\textwidth}
\begin{center}
\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.eps}
\end{center}
\end{minipage}
\\
\begin{minipage}{\textwidth}
\begin{center}
\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.eps}
\end{center}
\end{minipage}
\end{tabular}
\caption{PPI from each task that are found by our method but are missed by the homolog method. These are ones for which sequence-similar proteins were not found. (a) \textit{Hepatitis-C} human PPI and (b) \textit{Influenza}-human PPI.}
\label{fig:network}
\end{figure}

@@

\begin{table}[bht]
\begin{scriptsize}
\begin{center}
%\def\arraystretch{1.1}
\begin{tabular}{|c|c|}
\multicolumn{2}{c}{Gene Ontology terms for clusters from Figure \ref{fig:pca2}} \\ \hline
cluster 1 & mRNA (guanine-N7-)-methyltransferase activity; RNA-directed \\
(ebola only)& RNA polymerase activity; ATP binding \\ \hline
cluster 2 & host cell endoplasmic reticulum membrane; \\ 
(hepatitis \& & host cell lipid particle; induction by virus \\
influenza & of host autophagy; host cell mitochondrial membrane; \\
proteins) & suppression by virus of host STAT1 activity \\ \hline
cluster 3 (all) & fusion of virus membrane with host plasma membrane \\ \hline
\end{tabular}
\caption{GO terms (for process, function, component) enriched in the highlighted clusters from Figure \ref{fig:pca2}. FuncAssociate GO enrichment tool was used and terms with p-value < 0.001 are shown. The first column also indicates the composition of each cluster (i.e which task's proteins appear in the cluster).}
\label{goterms}
\end{center}
\end{scriptsize}
\end{table}

@@

\section{Results on HIV data}
We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus). The interactions data (obtained from PHISTO and NIAID databases) has $\approx$4000 human-HIV PPI across several strains, of which 1320 PPI are non-homologous. This experiment particularly tests the robustness of the multitask learning approaches by the addition of a relatively unrelated task. Table \ref{hivresults} shows the results.

\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting. Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2. Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of a somewhat dissimilar task such as HIV.}
\label{hivresults}
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|cccc}
\toprule
& \multicolumn{4}{c}{10 \% training } \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

\bibliographystyle{natbib}
\begin{scriptsize}
\bibliography{references}
\end{scriptsize}

>>>>

\documentclass[journal,12pt,onecolumn,draftclsnofoot,]{IEEEtran}

%\usepackage[top=1.3in, bottom=1.3in, left=1in, right=1in]{geometry}
\usepackage{floatrow}
\usepackage{graphicx} % more modern
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{subfigure} 
\usepackage{natbib}
\usepackage{hyperref}
%\usepackage[footnotesize]{caption}
\usepackage{multirow}
\usepackage{url}

\usepackage{etoolbox}
\makeatletter
\patchcmd{\@makecaption}
  {\scshape}
  {}
  {}
  {}
\makeatletter
\patchcmd{\@makecaption}
  {\\}
  {.\ }
  {}
  {}
\makeatother
\def\tablename{Table}
\renewcommand{\thetable}{\arabic{table}}

@@

\begin{document}

\title{Multitask matrix completion for learning protein interactions across diseases}

\author{Meghana Kshirsagar$^{1}$, Keerthiram Murugesan$^{2}$, Jaime G. Carbonell$^{2}$ and Judith Klein-Seetharaman$^{3}$}
%\address{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021\\
%$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
%$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK
%}

\begin{abstract}
\textbf{Motivation:}
Disease causing pathogens such as viruses, introduce their proteins into the host cells 
where they interact with the host's proteins enabling the virus to replicate inside the host.
These interactions between pathogen and host proteins are key to understanding infectious diseases.
Often multiple diseases involve phylogenetically related or biologically similar pathogens. Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}. \\
\textbf{Results:}
Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
We show how our model's parameters can be interpreted to reveal both general and specific
interaction-relevant characteristics of the viruses. \\
\textbf{Availability:}
Code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
\textbf{Contact:}
\href{mkshirsa@cs.cmu.edu}{meghana.ksagar@gmail.com}
\end{abstract}

@@

\begin{figure}
%trim option's parameter order: left bottom right top
%\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
\includegraphics[scale=0.34, trim = 0 0 0 0]{Kshirsagar.Fig_1.eps}
\caption{Principal component analysis (PCA) of virus proteins in the original feature space. The first two principal components are shown. Shape and color of the points indicates which virus that protein comes from.}%
\label{fig:pca1}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.25, trim = 1.1cm 0.9cm 0 0]{pca_bslmtl_with_clusters.eps}
\caption{Principal component analysis (PCA) of virus proteins in the projected subspace. The first two principal components are shown. Shape of the points indicates which virus that protein comes from. We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
\label{fig:pca2}
\end{figure}

@@

\section{Results}
The PPI data we obtain from PHISTO has homologous PPI within a task. Note that the presence of sequence similarity across virus proteins is a strong criterion for computational methods to work, however orthologs or paralogs of proteins within a task lead to redundancy in the data and some readers may find it harder to judge the contribution of the methods. We therefore present results on three sets: 
\begin{enumerate}
\item the original PPIs from PHISTO. AUC-PR is shown in Table \ref{resultsTable}
\item \textbf{Non-homologous partition}: after removal of all homologous PPI from the test data. Here we ensure that the test data does not contain homologs of PPI observed in the training data (this ensures there are no redundant examples). Further, every set of homologous PPIs in the test data is replaced with a single PPI. A BLAST sequence similarity cut-off of $1e^{-5}$ was used to find homologs. The results are in Table \ref{nohomresultsTable}.
\item \textbf{Homologous PPI partition} : only homologous PPI part of the test data. This is essentially the set of PPI removed in the set-(b) above. Table \ref{homresultsTable} shows these results. 
\end{enumerate}
Note that the partitions in (b) and (c) together comprise our test data. Also, the negatives in the test data are randomly split into the `no-homolog' and 'homolog' partitions while maintaining the number of negatives in each partition to be 100 times the number of positives.

@@

@@

@@

\begin{figure}
\begin{tabular}{cc}
\begin{minipage}{0.4\textwidth}
\begin{center}
\includegraphics[scale=0.25, trim = 2cm 1.5cm 1cm 1.5cm]{Kshirsagar.Fig_3a.eps}
\end{center}
\end{minipage}
&
\begin{minipage}{0.55\textwidth}
\begin{center}
\includegraphics[scale=0.35, trim = 2cm 1.5cm 1cm 1.5cm]{Kshirsagar.Fig_3b.eps}
\end{center}
\end{minipage}
\end{tabular}
\caption{PPI from each task that are found by our method but are missed by the homolog method. These are ones for which sequence-similar proteins were not found. (a) \textit{Hepatitis-C} human PPI and (b) \textit{Influenza}-human PPI.}
\label{fig:network}
\end{figure}

@@

@@

%\section{Results on HIV data}
%We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus). The interactions data (obtained from PHISTO and NIAID databases) has $\approx$4000 human-HIV PPI across several strains, of which 1320 PPI are non-homologous. This experiment particularly tests the robustness of the multitask learning approaches by the addition of a relatively unrelated task. Table \ref{hivresults} shows the results.

@@

\bibliographystyle{natbib}
\begin{normalsize}
\bibliography{references}
\end{normalsize}

\newpage

\begin{table}
\begin{footnotesize}
\caption{Tasks and their sizes (a column corresponds to one bipartite graph between human proteins and the virus indicated in the column header). More importantly, each task represents several strains of one virus species. We pool together PPI from all strains as the number of interactions per strain is very small and not representative. PPI data from different strains often involves homologous proteins (virus proteins or/and human proteins) as these were curated from different publications/studies involving the virus. The second row gives a conservative estimate of the number of non-homologous PPI. These were obtained by computing BLAST sequence similarity (see Section \ref{sec:phisto} for details). All pathogens are single stranded RNA viruses. The last row shows that each of our graphs is extremely sparse.} 
\label{datasets}
\begin{tabular}{c|r|r|r} 
\hline
Task $\rightarrow$ &{\textit{Influenza A} } &\textit{Hepatitis C} & {\textit{Ebola}}\\
& (28 strains) & (5 strains, many isolates)  & (3 strains) \\
number of HP PPIs$^a$ (positives) & 848 & 981 & 90 \\\hline
number of non-homologous PPIs & 274 & 270 & 75 \\ \hline
\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
total \# of virus proteins in & 542 & 163 & 38 \\
UniProtKB across strains$^b$ & & & \\ \hline
number of negatives & 84800 & 98100 & 9000 \\\hline
density of observed graph$^c$ (as \%) & .15 & .60 & .20 \\\hline
\multicolumn{4}{l}{\scriptsize{$^a$ HP PPI: host-pathogen protein protein interactions}}\\
\multicolumn{4}{l}{\scriptsize{$^c$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The denominator was computed considering all}} \\
\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} proteins from all strains of the virus and $\approx$ 100,000 human proteins.}}\\
\multicolumn{4}{l}{\scriptsize{$^b$ this is the number of `reviewed' proteins on UniProtKB from all strains in our data}} \\ \hline
\end{tabular}
\end{footnotesize}
\end{table}

\clearpage

\begin{table}[!t]
\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset. The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data. We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100). The standard deviation is also shown. The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold. The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task. This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
\label{resultsTable}
\begin{tabular}{c|ccccccc}
\toprule
& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
MMTL & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
Sparse+low-rank & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
MTPL & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
IMC & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
BSL-MTL & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
\end{tabular}
\end{table}

\clearpage

\begin{table}[!t]
\caption{AUC-PR on the \textbf{non-homologous PPI within each task}, in the 10\% setting. To get an estimate of the relative size of the non-homologous partition for each task, see Table \ref{datasets}. Only the most representative baselines are retained for this experiment. Our model performs better than the others on the non-homologous part of the test data, as it captures non-linear similarities among the interactions within a task and across tasks very well.}
\label{nohomresultsTable}
\begin{small}
%\def\arraystretch{1.2}
\begin{tabular}{c|ccc}
\toprule
\multicolumn{4}{c}{\textbf{10 \% training, 90\% test on Non-homologous PPI}} \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & 0.207 $\pm$.07 & 0.113 $\pm$.01 & 0.109 $\pm$ .01 \\
STL  & 0.210 $\pm$.14 & 0.474 $\pm$ .06 & 0.302 $\pm$.03 \\
Sparse+LR & 0.216 $\pm$.13 & 0.482 $\pm$.05 & 0.312 $\pm$.04 \\ 
BSL-MTL & \textbf{0.295} $\pm$.10 & \textbf{0.544} $\pm$.06 & \textbf{0.505} $\pm$.05 \\ \bottomrule
\end{tabular}
\end{small}
\end{table}

\clearpage

\begin{table}[!t]
\caption{AUC-PR on the \textbf{homologous PPI} within each task, in the 10\% setting. The Homolog baseline outperforms all on the \textit{Ebola} task and BSL-MTL beats all other methods on the other tasks. The Homolog method has access to a wider virus-human PPI graph involving several other viruses and is able to infer homologous PPI for \textit{Ebola} better (note that some of the homologous PPI are not redundant examples but involve proteins with very small regions of sequence similarity.}
\label{homresultsTable}
\begin{small}
%\def\arraystretch{1.2}
\begin{tabular}{c|ccc}
\toprule
\multicolumn{4}{c}{\textbf{10 \% training, test only Homologous PPI }} \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & \textbf{0.399} $\pm$.07 & 0.161 $\pm$.01 & 0.180 $\pm$ .01 \\
STL   & 0.192 $\pm$.15 & 0.504 $\pm$.10 & 0.305 $\pm$.03 \\
Sparse+LR & 0.184 $\pm$.13 & 0.512 $\pm$.10 & 0.315 $\pm$.04 \\ 
BSL-MTL & 0.130 $\pm$.10 & \textbf{0.815} $\pm$.02 & \textbf{0.475} $\pm$.02 \\ \bottomrule
\end{tabular}
\end{small}
\end{table}

\clearpage

\begin{table}[!t]
\begin{scriptsize}
\begin{center}
%\def\arraystretch{1.1}
\caption{GO terms (for process, function, component) enriched in the highlighted clusters from Figure \ref{fig:pca2}. FuncAssociate GO enrichment tool was used and terms with p-value $<$ 0.001 are shown. The first column also indicates the composition of each cluster (i.e which task's proteins appear in the cluster).}
\label{goterms}
\begin{tabular}{|c|c|}
\multicolumn{2}{c}{Gene Ontology terms for clusters from Figure \ref{fig:pca2}} \\ \hline
cluster 1 (ebola only) & mRNA (guanine-N7-)-methyltransferase activity; RNA-directed RNA polymerase activity; ATP binding \\ \hline
cluster 2 (hepatitis \& & host cell endoplasmic reticulum membrane; host cell lipid particle; induction by virus \\ 
influenza proteins) &  of host autophagy; host cell mitochondrial membrane; suppression by virus of host STAT1 activity \\ \hline
cluster 3 (all) & fusion of virus membrane with host plasma membrane \\ \hline
\end{tabular}
\end{center}
\end{scriptsize}
\end{table}

%\clearpage

%
%\begin{table}[!t]
%\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting. Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2. Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of a somewhat dissimilar task such as HIV.}
%\label{hivresults}
%\begin{small}
%%\def\arraystretch{1.2}
%\begin{tabular}{c|cccc}
%\toprule
%& \multicolumn{4}{c}{10 \% training } \\
%& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
%Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
%STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
%Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
%BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
%\end{tabular}
%\end{small}
%\end{table}
